A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2018-10-19 doc/large_strain.tex
add jacobian from isoparameter to real element in MF

The Newton-Raphson solution approach is adopted for the nonlinear problem considered here.
Given the current trial solution field $\overline{\gz u^h}$, the correction $\Delta \gz u^h$ field is obtained as the solution to the following system of equations (see \eqref{eq:weak_form} and \eqref{eq:tangent}):
\begin{align}
  A_{ij} \Delta u_j &= - F_i  \label{eq:linear_system} \\
  A_{ij} \equiv a(\gz N_i, \gz N_j) &=
  \int_{\mcl B_0}
  \underbrace{
  \overline{\grad^s} \gz N_i : J \boldsymbol{\mathcal{C}} : \overline{\grad^s} \, \gz N_j
  +
  \overline{\grad}\gz N_i :
  \left[
  \overline{\grad} \gz N_j \cdot
  \overline{\gz \tau}
  \right]
  }_{\displaystyle \tilde{a}(\gz N_i, \gz N_j)}
  \d V
  \label{eq:algebraic_tangent}
  \\
  F_i &=
  \int_{\mcl B_0} \overline{\gz \tau} : \grad^{s} \, \gz N_i \, \d V \,
  -
  \int_{\partial \mcl B_0^N} \overline{\gz T} \cdot \gz N_i \, \d S.
\end{align}
Here $\gz A$ is the discrete tangent matrix, $\gz F$ is the discrete gradient
and $\tilde{a}(\gz N_i, \gz N_j)$ is a representation for the integrand in the bilinear form for the trial solution field $\overline{\gz u^h}$.
Note that $\overline{\grad^s} \gz N_i$ and $\overline{\grad} \gz N_i$ are gradients of shape functions in the deformed configuration, whereas the integration is done in the reference (undeformed) configuration.

@@

        &\approx \sum_q \sum_j \tilde{a}(\gz N_i,\gz N_j)(\gz x_q) x_j W_q \\
        &= \sum_K \sum_q \sum_j \tilde{a}(\gz N_i,\gz N_j)(\gz \xi^K_q) x_j w_q

@@

where in the last step $\tilde{a}(\gz N_i,\gz N_j)(\gz \xi^K_q)$ is a representation for the evaluation of the bilinear form in one quadrature point $q$.

@@

 \tilde{a}(N_i,N_j)(\gz \xi^K_q) = N_i(\gz \xi^K_q)N_j(\gz \xi^K_q).

@@

where for quadrature points $q$ we can associate a multi-index $(q_1,q_2)$.

@@

Shape functions can also be expressed as product of $d$ one dimensional basis functions

@@

  y_{i} = y_{i_1\,i_2} & = \sum_{q_1} \sum_{q_2} \sum_{j_1} \sum_{j_2}

@@

  &=

@@

These four loops all have the same structure. We either keep the shape function fixed and iterate over the quadrature points in one spatial direction (in the first two sums) or keep the quadrature point and fixed and iterate over all the shape functions in one spatial direction (in the last two sums). Since each of these $2\cdot d$ loops has a complexity of $n$ and we perform them for $n^d$ values in quadrature points or values in degrees of freedoms simultaneously, we end up with an arithmetic complexity of $\mathcal{O}(dn^{d+1})$.

>>>>

The Newton-Raphson solution approach is adopted for the nonlinear problem considered here.
Given the current trial solution field $\overline{\gz u^h}$, the correction $\Delta \gz u^h$ field is obtained as the solution to the following system of equations (see \eqref{eq:weak_form} and \eqref{eq:tangent}):
\begin{align}
  A_{ij} \Delta u_j &= - F_i  \label{eq:linear_system} \\
  A_{ij} \equiv a(\gz N_i, \gz N_j) &=
  \int_{\mcl B_0}
  \underbrace{
  \overline{\grad^s} \gz N_i : J \boldsymbol{\mathcal{C}} : \overline{\grad^s} \, \gz N_j
  +
  \overline{\grad}\gz N_i :
  \left[
  \overline{\grad} \gz N_j \cdot
  \overline{\gz \tau}
  \right]
  }_{\displaystyle \tilde{a}(\gz N_i, \gz N_j)}
  \d V
  \label{eq:algebraic_tangent}
  \\
  F_i &=
  \int_{\mcl B_0} \overline{\gz \tau} : \grad^{s} \, \gz N_i \, \d V \,
  -
  \int_{\partial \mcl B_0^N} \overline{\gz T} \cdot \gz N_i \, \d S.
\end{align}
Here $\gz A$ is the discrete tangent matrix, $\gz F$ is the discrete gradient
and $\tilde{a}(\gz N_i, \gz N_j)$ is a representation for the integrand in the bilinear form $a(\gz N_i, \gz N_j)$ for the trial solution field $\overline{\gz u^h}$.
Note that $\overline{\grad^s} \gz N_i$ and $\overline{\grad} \gz N_i$ are gradients of shape functions in the deformed configuration, whereas the integration is done in the reference (undeformed) configuration.

@@

        &\approx \sum_q \sum_j \tilde{a}(\gz N_i,\gz N_j)(\gz x_q) x_j w_q \\
        &= \sum_K \sum_q \sum_j \tilde{a}(\gz N_i,\gz N_j)(\gz \xi_q) x_j w_q J^K

@@

where in the last step $\tilde{a}(\gz N_i,\gz N_j)(\gz \xi_q)$ is a representation for the evaluation of the bilinear form in one quadrature point on isoparametric element $\gz \xi_q$ and $J^K$ is the Jacobian of the mapping from the isoparametric to the element $K$.

@@

 \tilde{a}(N_i,N_j)(\gz \xi_q) = N_i(\gz \xi_q)N_j(\gz \xi_q).

@@

where for each quadrature points $q$ we can associate a multi-index $(q_1,q_2)$.

@@

Shape functions can also be expressed as product of one dimensional basis functions

@@

  y_{i} = y_{i_1\,i_2} & = J^K \sum_{q_1} \sum_{q_2} \sum_{j_1} \sum_{j_2}

@@

  &= J^K

@@

These four loops all have the same structure.
We either keep the shape function fixed and iterate over the quadrature points in one spatial direction (in the first two sums) or keep the quadrature point fixed and iterate over all the shape functions in one spatial direction (in the last two sums).
Since each of these $2\times d$ loops has a complexity of $n$ and we perform them for $n^d$ values in quadrature points or values in degrees of freedoms simultaneously, we end up with an arithmetic complexity of $\mathcal{O}(dn^{d+1})$.
